Applying State-free Reversible VAMPNets and Markov State Models to Learn Dynamics of DNA Oligonucleotides

### 2021-01-27 manuscript.tex
Updates from Overleaf

Despite rapid advances in DNA nanotechnology and a robust understanding of the associated thermodynamics, the sequence-dependent mechanisms of DNA hybridization are not fully understood. In this work, we investigate these dynamics by performing equilibrium coarse-grained simulations of oligonucleotide sequences with varied G:C placement. We employ State-Free Reversible VAMPnets to directly learn the slowest dynamical modes of each sequence and optimize Markov State Model (MSM) construction. Furthermore, we perform elevated temperature simulations to recapitulate temperature-jump IR and FTIR data collected on the oligonucleotides. For repetitive sequences, we find a spectrum of slow dynamics associated with out-of-register base pairing and kinetically relevant transitions between these states. In contrast, G:C pairs near the center of the duplex induce more pronounced fraying dynamics through which hybridization and dissociation are facilitated. In both cases, these mechanisms deviate from an “all-or-nothing” hybridization model. Our computational predictions show agreement with experiments and provide new fundamental understanding of the sequence-dependent kinetics and mechanisms of DNA hybridization.

>>>>

Despite rapid advances in DNA nanotechnology and a robust understanding of the associated thermodynamics, the sequence-dependent mechanisms of DNA hybridization are not fully understood. In this work, we investigate these dynamics by performing equilibrium coarse-grained simulations of oligonucleotide sequences with varied G:C placement. We employ State-Free Reversible VAMPnets to directly learn the slowest dynamical modes of each sequence and optimize Markov State Model construction. Furthermore, we perform elevated temperature simulations to recapitulate temperature-jump IR and FTIR data collected on the oligonucleotides. For repetitive sequences, we find a spectrum of slow dynamics associated with out-of-register base pairing and kinetically relevant transitions between these states. In contrast, G:C pairs near the center of the duplex induce more pronounced fraying dynamics through which hybridization and dissociation are facilitated. In both cases, these mechanisms deviate from an “all-or-nothing” hybridization model. Our computational predictions show agreement with experiments and provide new fundamental understanding of the sequence-dependent kinetics and mechanisms of DNA hybridization.
